Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-02-04 submission5/core.tex
Resolved more comments

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}. \sna{Nowhere clearly enough. We need to make it clear in section 6
  that Nick and Andre are included there---only aiming at caller-side,
  not callee-side. The discussion of CHERI work should touch on this as well.}
The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination. Our model is
  modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate a tag-based enforcement mechanism, \emph{Lazy
  Tagging and Clearing}, via property-based random testing, find that
it falls short, and propose and validate a fix.
\end{itemize}

@@

We define stack safety properties in
terms of traces of {\em observable events} rather than internal machine states.
This is particularly important for
specifying lazy enforcement, which catches bad behaviors after they occur but
before they can affect the trace.
Observable events might include specific function calls (e.g., system
calls that perform I/O), writes to special addresses representing
memory-mapped regions, etc.
%The nature of events is a parameter of the framework, and can be specialized
%to any notion of observable behavior.

%% \paragraph*{Variants and Non-interference}

To characterize situations in which values should be kept secret or prevented
from influencing future execution, we borrow the idea of \emph{variant} states
from the theory of non-interference.  For example, to show that certain locations
are kept secret, it suffices to vary their contents them and check that the observable
output does not change.

%Non-interference is a way of expressing
%the flow of information through the system. If we begin with a state \(\mach\)
%and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
%then execute both \(\mach\) and \(\nach\), any differences in the outputs they
%produce leaks information about \(\component\).

@@

%The security principals in our account of stack safety are function
%activations: the context tracks, for the active function and all pending
%functions, their {\em view} of the security state of the system.

@@

\paragraph*{Security Semantics and Property Structure}
A security semantics extends a machine
with additional context about the identities of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the behavior of the real machine. The security context
evolves dynamically through the execution of security-relevant operations,
which include events like calls, returns, and frame manipulation.
Our security properties are phrased in terms of this context, often as predicates
on future states (``when control returns to the current function...'')
or as relations on traces of future execution
(hyper-properties).\bcp{This is the first we've mentioned
  hyperproperties.  CSF readers may be surprised.}
\sna{Surprised that we have hyperproperties,
  or that we made it this far without mentioning them?}
\bcp{The latter!}

Security-relevant operations typically correspond to
single underlying machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V architecture that we use in our
examples and tests,
calls and returns are conventionally performed using the {\tt jal}
(``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, but these
instructions might also be used for other things. Thus we identify
security-relevant operations by labeling selected instructions.

@@

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. A core set of operations covering calls, returns, and local
memory is introduced in the example in \cref{sec:example}
and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
The remaining operations are needed for the capability-based model in \cref{app:ptr}. 
%\bcp{I was confused by the table because I looked at
% the figure before I read that sentence!}
%Those that we explore in detail in \cref{sec:example,sec:formal}
%are shown in white: calls, returns, and the allocation and deallocation of local memory.
%In \cref{sec:extensions}, we cover additional operations and and parameters
%to support various forms of memory sharing \ifexceptions , exceptions, \fi
%and tail-call elimination; these are shown in light gray. These are also
%the operations we test in \cref{sec:testing}.

\bcp{As a reader, I'm feeling a bit overwhelmed at this point.
  Everything feels fluid and flexible---''It could be like this, but
  it could also be another way if you want...''  I know we're past the
  point of doing more than tweaking the way we talk about things, but
  if there are any tweaks that reduce the sense of flexibility, let's
  do them!}

\sna{Agreed. The point of the operations was originally that we shouldn't
  have to think about all of these things when presenting the formalism,
  not that we should open with all the different ways we can think about them.}

\apt{Hopefully better now.}

@@

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over: (1) the details of  the target machine
architecture and ABI, (2) the set of security-relevant operations and their effects on
the security context, (3) the set of observable events, and (4) a notion of value compatibility.
%They can be extended to a more sophisticated machine model by expanding the set of
%security-relevant events and defining
\apt{Any more? Give a more listlike presentation?
  Need to clarify exactly what the parameters are, both in the initial
  presentation and in the various extensions, including the appendix.}
\sna{Need to come back to this.}

@@

Using this mechanism, a call-by-value argument passed on the stack at an \(\SP\)-relative offset
is specified by the triple \((\SP, \mathit{off}, \mathit{sz})\).
In this case, only the immediate callee gains access to the argument location.
A C++-style call-by-reference argument % where the reference is
passed in \(\reg\)
is instead specified by \((\reg, 0, \mathit{sz})\), which allows
the reference to be passed further down the stack.
\apt{What if the reference is itself passed on the stack?}
\sna{Ugh. We could make \(\reg\) an arbitrary element, but that seems like it's getting
  dangerously close to indirection.}

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}.
%, which we survey in Sections~\ref{sec:intro} and~\ref{sec:future}.
% Among these techniques,
%
Many enforcement techniques focus purely on \(\wbcf\);
others combine this with some degree of memory protection,
chiefly focusing on integrity.
%
%The micro-policies of
Roessler and
DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
range of safeguards including lazy enforcement, and suitable to formulate
and validate strong security guarantees.

To our knowledge, the only other line of work that aims to rigorously characterize the
security of the stack is the Cerise family of program logics
\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
Most of those define experimental types of hardware capabilities to achieve
efficient implementations.
%
These works define secure calling conventions for capability machines and
overlay semantics that internalize those conventions and their related stack
safety properties, phrased in terms of logical relations instead of trace
properties.
%
Originally, they describe ``local state
encapsulation''~\cite{SkorstengaardSTKJFP} in terms of integrity (but it is
equivalent to % the combination of
\(\clri\) {\em and} \(\clrc\)). Their latest
paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
in this paper to extend their formalism to include confidentiality. When
checking if our properties applied to their old calling convention, they noted that
it did not enforce \(\clec\), and made sure that their new version
would in addition to building it into their formalism
\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
choice to explicitly state properties in security terms: specifying security
% mechanisms
is hard, and when the spec takes the form of a ``correct by
construction'' machine, it is easy to neglect a non-obvious security
requirement.

\sna{See commented out tex for content that could be massaged to fit in here,
  depending on whether we're interested in delving deeper.}\bcp{If
  there's space, some of this could be resuscitated.}
\apt{It would be very good to do this, because reviewers will want a careful comparison with this work,
  but I don't feel quite confident enough to do it (especially after my unfortunate attempt last time)!}
%For instance, the stack-safety specification from Georges et al.
%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
%a callee should always return to its caller. But, in the presence of tail-call elimination, a
%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).

@@

\apt{Sbould admit up front that this extension requires a fundamental change in the
  security context.}

What if we want to express a finer-grained notion of safety, in which
stack objects are protected unless the function that owns them intentionally
passes a pointer to them? This can be thought of as a {\it capability}-based
notion of security. Capabilities are unforgeable tokens that grant access to
a region of memory, typically corresponding to valid pointers to that region.
So, in order to express such a property, we need our machine to carry some notion
of {\it pointer provenance}---a distinction between a pointer that is intended to
point to a given object, and non-pointer integers as well as pointers to other objects.

@@

Int the resulting property, once an object is sealed (because its
capability has not been passed to a callee), subsequent nested calls can never unseal it.
On the other hand, an object that is passed via a pointer may be passed on indefinitely.

>>>>

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
  of properties: confidentiality and integrity for callee and caller,
  plus well-bracketed control-flow.
  The properties are parameterized over a notion of
  external observation, allowing them to characterize lazy enforcement
  mechanisms.
\item We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination. Our model is
  modular enough that handling these features is straightforward.
\item We validate a tag-based enforcement mechanism, \emph{Lazy
  Tagging and Clearing}, via property-based random testing, find that
it falls short, and propose and validate a fix.
\end{itemize}

@@

We define stack safety properties in
terms of traces of {\em observable events} rather than internal machine states.
This is particularly important for
specifying lazy enforcement, which catches bad behaviors after they occur but
before they can affect the trace.
Observable events might include specific function calls (e.g., system
calls that perform I/O), writes to special addresses representing
memory-mapped regions, etc.

To characterize situations in which values should be kept secret or prevented
from influencing future execution, we borrow the idea of \emph{variant} states
from the theory of non-interference.  For example, to show that certain locations
are kept secret, it suffices to vary their contents and check that the observable
output does not change.

@@

@@

\paragraph*{Security Semantics and Property Structure}
A security semantics extends a machine
with additional context about the identities of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the behavior of the real machine. The security context
evolves dynamically through the execution of security-relevant operations,
which include calls, returns, and frame manipulation.
Our security properties are phrased in terms of this context, often as predicates
on future states (``when control returns to the current function...'')
or as relations on traces of future execution
(hyper-properties).

Security-relevant operations abstract over the implementation details of the
actions they take. In this paper, each operation corresponds to a
single underlying machine instruction. One instruction may perform multiple
operations. We assume that a compiler or other trusted source has provided
labels to disambiguate instructions that have multiple purposes. For instance,
in the tagged RISC-V architecture
that we use in our examples and tests,
calls and returns are conventionally performed using the {\tt jal}
(``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, but these
instructions might also be used for other things.

@@

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. A core set of operations covering calls, returns, and local
memory is introduced in the example in \cref{sec:example}
and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
The remaining operations are needed for the capability-based model in \cref{app:ptr}.

@@

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over: (i) the details of  the target machine
architecture and ABI, (ii) the set of security-relevant operations and their effects on
the security context, (iii) the set of observable events, and (iv) a notion of value compatibility.

@@

Using this mechanism, a call-by-value argument passed on the stack at an \(\SP\)-relative offset
is specified by the triple \((\SP, \mathit{off}, \mathit{sz})\).
In this case, only the immediate callee gains access to the argument location.
A C++-style call-by-reference argument % where the reference is
passed in \(\reg\)
is instead specified by \((\reg, 0, \mathit{sz})\), which allows
the reference to be passed further down the stack.

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}.
%, which we survey in Sections~\ref{sec:intro} and~\ref{sec:future}.
% Among these techniques,
%
Many enforcement techniques focus purely on \(\wbcf\);
others combine this with some degree of memory protection,
chiefly focusing on integrity.
%
Roessler and DeHon's {\it Depth Isolation} and {\it Lazy Tagging and Clearing}
\cite{DBLP:conf/sp/RoesslerD18} both offer protections corresponding to
\(\wbcf\), \(\clri\), and \(\clrc\), though they do not give a formal description
of this. They are generally not concerned with protecting callees.

To our knowledge, the only other line of work that aims to rigorously characterize the
security of the stack is the Cerise family of program logics
\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
Most of those define experimental types of hardware capabilities to achieve
efficient implementations.
%
These works define secure calling conventions for capability machines and
overlay semantics that internalize those conventions and their related stack
safety properties, phrased in terms of logical relations instead of trace
properties.
%
Originally, they describe ``local state
encapsulation''~\cite{SkorstengaardSTKJFP} in terms of integrity (but it is
equivalent to % the combination of
\(\clri\) {\em and} \(\clrc\)). Their latest
paper \cite{Georges22:TempsDesCerises} was inspired by the properties presented
in this paper to extend their formalism to include confidentiality. When
checking if our properties applied to their old calling convention, they noted that
it did not enforce \(\clec\), and made sure that their new version
would in addition to building it into their formalism
\cite{Georges22:personalcommunication}. This demonstrates the benefit of our
choice to explicitly state properties in security terms: specifying security
% mechanisms
is hard, and when the spec takes the form of a ``correct by
construction'' machine, it is easy to neglect a non-obvious security
requirement.
%
%For instance, the stack-safety specification from Georges et al.
%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
%a callee should always return to its caller. But, in the presence of tail-call elimination, a
%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).

@@

What if we want to express a finer-grained notion of safety, in which
stack objects are protected unless the function that owns them intentionally
passes a pointer to them? This can be thought of as a {\it capability}-based
notion of security. Capabilities are unforgeable tokens that grant access to
a region of memory, typically corresponding to valid pointers to that region.
So, in order to express such a property, we need our machine to carry some notion
of {\it pointer provenance}---a distinction between a pointer that is intended to
point to a given object, and non-pointer integers as well as pointers to other objects.
This constitutes a substantial addition to the security context.

@@

In the resulting property, once an object is sealed (because its
capability has not been passed to a callee), subsequent nested calls can never unseal it.
On the other hand, an object that is passed via a pointer may be passed on indefinitely.
